---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# erglm

<!-- badges: start -->
[![R-CMD-check](https://github.com/djnavarro/erglm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/djnavarro/erglm/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/djnavarro/erglm/graph/badge.svg)](https://app.codecov.io/gh/djnavarro/erglm)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/erglm)](https://CRAN.R-project.org/package=erglm)
<!-- badges: end -->

Provides estimation tools for exposure-response models based on `glm()`. It is mostly intended as a convenience package: the core tools are wrappers around `glm()`, tested and supported for binomial, poisson, gaussian, and gamma families. For plotting exposure-response models (including those fitted with erglm), see the companion package [erplots](https://erplots.djnavarro.net), which supplies a model-agnostic mini-language for building exposure-response plots. For an analogous approach to time-to-event models see [ertte](https://ertte.djnavarro.net), and for Emax regression models see [emaxnls](https://emaxnls.djnavarro.net).

## Installation

You can install the latest CRAN release of erglm with this:

``` r
install.packages("erglm")
```

Alternatively you can install the development version of erglm like so:

``` r
pak::pak("djnavarro/erglm")
```

## Models

```{r}
#| label: erglm-model

library(erglm)

head(erglm_data)

mod <- erglm_model(ae1 ~ aucss, erglm_data, family = binomial())
mod
```

## Stepwise covariate modelling

```{r}
mod1 <- erglm_model(ae1 ~ aucss + sex + dose, erglm_data, family = binomial())
mod2 <- erglm_scm_backward(mod1, candidates = c("sex", "dose"))
erglm_scm_history(mod2)
```

## Simulation

```{r}
#| label: erglm-sim
mod <- erglm_model(ae1 ~ aucss + sex, erglm_data, family = binomial())
sim <- simulate(mod, nsim = 5, seed = 1234)
head(sim)
```

To visualise the simulations against the observed data (e.g. as a VPC-style
plot) see erplots' `er_vpc()` mini-grammar -- specifically
`erplots::er_vpc_add_simulated()`, which can build its own simulated
replicates directly from a fitted `erglm_model`
(`er_vpc_add_simulated(model = mod)`).

