---
title: "Getting Started"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

erglm provides estimation tools for exposure-response models based on
`glm()`. It's mostly a convenience package: the core tools are thin
wrappers around `glm()` and its usual machinery, tested and supported for
binomial, poisson, gaussian, and gamma families. This article is a quick
tour of the main pieces; the other articles go into more depth on each.

```{r setup}
library(erglm)
library(tibble)
```

## Example data

The package ships with a synthetic dataset, `erglm_data`, used throughout
the documentation. It has an exposure metric (`aucss`), several covariates
(`sex`, `age`, `weight`, `dose`, ...), and response columns for each
supported family: binary (`ae1`, `ae2`), count (`ae_count`), continuous
(`biomarker_change`), and positive right-skewed continuous (`ae_duration`):

```{r data}
erglm_data
```

## Fitting a model

`erglm_model()` fits the model -- it takes the same `formula`/`data`
arguments as `glm()`, plus a `family` argument (defaulting to
`gaussian()`, matching `glm()`'s own default):

```{r model}
mod <- erglm_model(ae1 ~ aucss + sex, erglm_data, family = binomial())
mod
```

## Prediction

`erglm_predict()` produces predictions with confidence intervals, on both
the link and response scales, as a tidy data frame:

```{r predict}
mod |>
  erglm_predict(newdata = tibble(aucss = seq(0, 3000, by = 500), sex = "Female"))
```

## Choosing covariates

When there are several candidate covariates, `erglm_scm_forward()` and
`erglm_scm_backward()` automate the process of deciding which belong in
the model, via stepwise addition/elimination based on significance
testing:

```{r scm}
erglm_scm_forward(mod, candidates = c("dose", "weight", "age"), seed = 1024)
```

See the ["Stepwise covariate modelling"](scm.html) article for a full
treatment, including the forward/backward workflow and the audit log
(`erglm_scm_history()`).

## Simulation

`simulate()` generates replicate datasets from a fitted model, and
`erglm_vpc_sim()` reshapes those replicates into a data set ready for a
visual predictive check:

```{r simulate}
erglm_vpc_sim(mod, nsim = 5, seed = 2048)
```

See the ["Simulation"](simulate.html) article for details, including the
lower-level `erglm_fun()` building block.

## Working with fitted models

An object returned by `erglm_model()` is a genuine `glm` object -- it has
class `c("erglm_model", "glm", "lm")` -- so all of the standard `glm`/`lm`
methods (`summary()`, `predict()`, `confint()`, `AIC()`, `anova()`, and so
on) work on it directly, with no erglm-specific replacement needed. See the
["Using base R model methods"](methods.html) article for a worked-through
tour of these.

## Where to next

- ["Modelling"](model.html) -- more on fitting and prediction, including
  the other `glm()` families erglm supports.
- ["Stepwise covariate modelling"](scm.html) -- the forward/backward SCM
  workflow in full.
- ["Using base R model methods"](methods.html) -- the `glm`/`lm` methods
  that come for free.
- ["Simulation"](simulate.html) -- `simulate()`, `erglm_fun()`, and
  `erglm_vpc_sim()`.
- the companion [erplots](https://github.com/djnavarro/erplots) package,
  for visualising exposure-response models fitted with erglm.
