---
title: "Using base R model methods"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(erglm)
```

An object returned by `erglm_model()` is a genuine `glm` object -- it has
class `c("erglm_model", "glm", "lm")`, and the `erglm_model` class only adds
a little extra metadata on top. This means none of the standard R methods
for working with fitted models are erglm-specific: they're the same ones
you'd use for a plain `glm()` fit, and they work here without any
modification. This article is a short reference for the ones that come up
most often when working with exposure-response models.

```{r model}
mod <- erglm_model(ae1 ~ aucss + dose + sex, erglm_data, family = binomial())
```

## Model summary

`summary()` gives the usual coefficient table, standard errors, and
dispersion information:

```{r summary}
summary(mod)
```

## Coefficients and the variance-covariance matrix

The fitted coefficients and their variance-covariance matrix are available
via `coef()` and `vcov()`, exactly as for any `glm`:

```{r coef}
coef(mod)
```

```{r vcov}
vcov(mod)
```

Confidence intervals for individual coefficients can be obtained from
`confint()` (profile-likelihood based, so slightly slower but generally
preferred over a Wald interval for `glm` models):

```{r confint}
confint(mod)
```

## Predictions

Base R's `predict()` works directly on an erglm model. By default it
returns predictions on the link scale; use `type = "response"` for the
response scale:

```{r predict}
predict(mod, newdata = erglm_data[1:5, ], type = "response")
```

`predict()` can also return standard errors (`se.fit = TRUE`), which is
the basis for erglm's own [erglm_predict()] -- that function is a thin,
opinionated wrapper that back-transforms the link-scale standard errors
into a confidence interval on the response scale and returns everything as
a tidy data frame bound to `newdata`:

```{r erglm_predict}
erglm_predict(mod, newdata = erglm_data[1:5, ])
```

Use whichever is more convenient: `predict()` for a quick point estimate,
or `erglm_predict()` when you want interval bounds without computing them
by hand.

## Model comparison

`AIC()`, `BIC()`, and `logLik()` all work as usual, which is convenient for
comparing candidate models outside of erglm's own stepwise covariate
modelling (`erglm_scm_forward()`/`erglm_scm_backward()`):

```{r aic-bic}
mod_no_sex <- erglm_model(ae1 ~ aucss + dose, erglm_data, family = binomial())

AIC(mod, mod_no_sex)
BIC(mod, mod_no_sex)
```

For nested models, `anova()` gives a likelihood-ratio (or F, for families
with estimated dispersion) test directly -- this is exactly the machinery
`erglm_scm_forward()`/`erglm_scm_backward()` use internally, exposed here
for one-off comparisons:

```{r anova}
anova(mod_no_sex, mod, test = "Chisq")
```

## A note on diagnostic plots

Base R's `plot.lm()` diagnostic plots (`plot(mod, which = 1:4)`) also work,
though some panels (e.g. leverage) are more informative for continuous
responses than for binary ones. erglm deliberately doesn't provide its own
diagnostic plotting -- see the companion
[erplots](https://github.com/djnavarro/erplots) package for
exposure-response-specific visualisations.
