---
title: "Stepwise covariate modelling"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(digits = 4)
```

```{r setup, message = FALSE}
library(erglm)
library(tibble)
```

Once you can fit a single exposure-response model, the next practical question is usually *which covariates belong in the model*. When there are several candidates, testing every combination by hand is tedious and easy to get wrong. `erglm_scm_forward()` and `erglm_scm_backward()` automate this with **stepwise covariate modelling** (SCM): a forward-addition step that greedily adds the most helpful covariates, a backward-elimination step that prunes terms that no longer earn their place, and a complete history of every model considered along the way that serves as an audit log for the procedure.

## The building blocks

At the lowest level, covariate modelling is just adding or removing a single term from a fitted model. Two exported helpers do exactly that, `erglm_add_term()` and `erglm_remove_term()`, each taking a one-sided formula naming the term and returning a refitted model:

```{r building-blocks}
base_mod <- erglm_model(ae1 ~ aucss, erglm_data, family = binomial())

erglm_add_term(base_mod, ~ sex)
```

The stepwise functions are built on top of these: they repeatedly propose add/remove moves, compare each candidate against the current model by significance testing (`anova()`), and keep the best one. You rarely need to call `erglm_add_term()`/`erglm_remove_term()` directly, but it's worth knowing they're what the automated search is doing under the hood.

## Setting up the search

Two ingredients are needed to drive an SCM run: a **base model** to start from, and a character vector of **candidate covariates** to consider. Unlike some covariate-modelling frameworks, candidates here don't need to be tied to a particular structural parameter -- they're just term names that could be added to (or removed from) the right-hand side of the model formula:

```{r candidates}
candidates <- c("sex", "dose", "weight", "age")
```

## Forward addition

Every candidate not already in the model is added in turn and compared against the current model; the term with the smallest $p$-value is retained, provided that $p$-value is below `threshold`. The step repeats until no remaining candidate clears the bar:

```{r forward}
fwd_mod <- erglm_scm_forward(base_mod, candidates, threshold = 0.01, seed = 3425)
fwd_mod$formula
```

None of the four candidates clear the default $0.01$ threshold here -- inspecting the history shows why:

```{r forward-history}
erglm_scm_history(fwd_mod)[, c("term_tested", "term_p_value", "model_updated")]
```

`sex` does have a genuine (if modest) effect on `ae1` in the data-generating process behind `erglm_data`, but once `aucss` is already in the model that effect isn't strong enough, at this sample size, to clear a $0.01$ threshold -- so the search correctly leaves it out rather than adding a covariate that doesn't earn its place. This is a feature, not a limitation: SCM is exactly the tool that should stop you from including underpowered terms.

## Backward elimination

Backward elimination works in reverse: each term currently in the model is dropped in turn, and the term with the *largest* $p$-value is removed if that $p$-value exceeds `threshold`. To see it doing real work, start from a saturated model that already contains every candidate:

```{r backward}
full_mod <- erglm_model(ae1 ~ aucss + sex + dose + weight + age, erglm_data, family = binomial())
bwd_mod <- erglm_scm_backward(full_mod, candidates, threshold = 0.001, seed = 9821)
bwd_mod$formula
```

All four candidates are pruned, leaving only `aucss` -- matching what forward addition already told us: none of `sex`, `dose`, `weight`, or `age` earns a place in the model at these thresholds. The full audit log shows the iteration-by-iteration elimination:

```{r backward-history}
print(
  erglm_scm_history(bwd_mod)[, c("iteration", "term_tested", "term_p_value", "model_updated")],
  n = Inf
)
```

Each iteration removes the single worst-performing term (`model_updated == 1`); the next iteration re-tests everything still in the model, since removing one term can change the others' apparent significance.

**Why the thresholds differ.** It's standard practice to make forward addition more permissive than backward elimination -- $0.01$ to add but $0.001$ to retain. A term added on the looser forward criterion must then survive the stricter backward criterion, which guards against terms that only looked useful in the presence of others. Keeping the backward threshold at or below the forward threshold also prevents the procedure from cycling (adding and removing the same term forever).

**Reproducibility.** Within each step, candidate terms are tested in a random order, so results can depend on the state of the random number generator; passing a `seed` makes a run reproducible. If `seed` is omitted, a random one is chosen and reported via a message.

## Forward addition piped to backward elimination

The typical workflow is a **forward-backward** run: forward addition to build the model up, immediately followed by backward elimination to prune it. Both functions take a fitted model as their first argument and return a fitted model, so they compose naturally with the pipe:

```{r pipeline}
final_mod <- base_mod |>
  erglm_scm_forward(candidates, threshold = 0.01, seed = 3425) |>
  erglm_scm_backward(candidates, threshold = 0.001, seed = 9821)

final_mod$formula
```

Since forward addition already found nothing worth adding, backward elimination has nothing left to prune, and the pipeline settles on the base model. `final_mod` is an ordinary fitted model object, so all the usual methods apply -- `summary()`, `confint()`, `predict()`, and so on all behave normally.

## Other `glm()` families

Nothing about the SCM interface changes across `glm()` families. The same candidates, applied to the Poisson-distributed `ae_count` response, tell an analogous story -- `sex` has a borderline effect (visible in the full-model summary as $p \approx 0.07$) that still doesn't clear the default forward threshold:

```{r poisson}
base_pois <- erglm_model(ae_count ~ aucss, erglm_data, family = poisson())
fwd_pois <- erglm_scm_forward(base_pois, candidates, threshold = 0.01, seed = 3425)
fwd_pois$formula
```

Internally, `erglm_scm_forward()`/`erglm_scm_backward()` pick a likelihood-ratio chi-squared test for families with known dispersion (`binomial`, `poisson`, as used above) and an F-test for families with an estimated dispersion parameter (`gaussian`, `Gamma`), matching `stats::anova()`'s own `test` argument -- set the `test` argument explicitly to override this.

## Notes and caveats

- **Selection criterion.** The current implementation selects on $p$-values only. The history records `model_aic` and `model_bic` for every candidate, so you can audit the search against information criteria even though they aren't used to drive it.
- **Threshold choice.** The forward and backward thresholds are the main levers you control. Stricter thresholds yield sparser models.
- **Greediness.** Stepwise search is greedy and isn't guaranteed to find the globally best subset of covariates -- and, as this article's example shows, it may correctly find that *no* candidate covariates belong in the model at a given sample size and threshold. The audit log is valuable precisely because it makes the path the search took transparent and reproducible.
