---
title: "Simulation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is the simulation article

```{r setup}
library(erglm)
```

## Visual predictive checks

`erglm_vpc_sim()` draws simulated response values that reflect both parameter uncertainty (by resampling coefficients from the model's asymptotic sampling distribution) and, family-appropriately, residual noise around the resulting expectation:

```{r vpc-1}
mod <- erglm_model(ae2 ~ aucss + sex, erglm_data, family = binomial())
sim <- erglm_vpc_sim(mod, nsim = 20, seed = 1234)
sim
```

The noise model is family-specific: Bernoulli draws for `binomial`, Poisson draws for `poisson`, normal draws for `gaussian`, and gamma draws for `Gamma`. For example, with a Poisson response:

```{r vpc-2}
mod_pois <- erglm_model(ae_count ~ aucss + sex, erglm_data, family = poisson())
sim_pois <- erglm_vpc_sim(mod_pois, nsim = 20, seed = 1234)
sim_pois
```

Other `glm()` families aren't currently supported by `erglm_vpc_sim()` and will raise an informative error rather than silently falling back to an expectation-only draw.
